\tit

### 2021-02-28 paper.tex
misc tidying

%% \acmtrue

@@

Since {\AB{u} \AF ⊙ \AIC{[]} \AD ≡ \AB{u}}, it follows that\footnote{For functions \AB{f} and \AB{g}, {\AB{f} \AF{≗} \AB{g}} is extensional equality, i.e., {\AS ∀ \AB x \AS → \AB{f} \AB{x} \AD{≡} \AB{g} \AB{x}}.}
\agda{ν∘𝒟}
Moreover, the function \AF{𝒟} distributes over \AIC{[]} and \AF{\_⊙\_}:\out{\footnote{In algebraic terms, argument-flipped \AF{𝒟} is a monoid homomorphism from lists to language endofunctions.}}
\agda{𝒟[]⊙}
These facts suggest that \AB{𝒟} can be computed one list element at a time via a more specialized operation, as is indeed the case:\footnote{Repeated \AF{δ} application is expressed as a standard left fold\out{ \stdlibCitep{Data.List}}:
\agda{foldl}\vspace{-2ex}
}
\agda{δ}
\vspace{-4ex}
\agda{𝒟foldl}
Each use of \AF{δ} thus takes us one step closer to reducing general language membership to nullability.
%% In other words,%
%% \agda{ν∘foldlδ}
This simple observation is the semantic heart of the syntactic technique of \emph{derivatives of regular expressions} as used for efficient recognition of regular languages \citep{Brzozowski64}, later extended to parsing general context-free languages \citep{Might2010YaccID}.
Lemmas \AF{ν∘𝒟} and \AF{𝒟foldl} liberate this technique from the assumption that languages are represented \emph{symbolically}, by some form of grammar (e.g., regular or context-free), inviting other representations as we will see in \secref{Automatic Differentiation}.

@@

Decidable parsing (via {\AF{⟦\_⟧}}) relies only on \AF{ν} and \AF{δ}, which are defined in \figrefdef{symbolic-defs}{Symbolic differentiation (column-major/patterns)}{\agda{defs}}, as systematically derived from the lemmas of \figref{nu-delta-lemmas}.\out{\notefoot{Describe this derivation.}}
These definitions are meant to be read in ``column-major'' order, i.e., each column is one function definition.
Correctness is guaranteed by the types of \AF{ν} and \AF{δ} and so is proved automatically by type-checking \figref{symbolic-defs}.
(We could thus use \emph{any} definitions of \AF{ν} and \AF{δ} that type- and termination-check, although not many definitions would.)

>>>>

\acmtrue

@@

Since {\AB{u} \AF ⊙ \AIC{[]} \AD ≡ \AB{u}}, it follows that\footnote{For functions \AB{f} and \AB{g}, {\AB{f} \AF{≗} \AB{g}} is extensional equality, i.e., {\AS ∀ \AB x \AS → \AB{f} \AB{x} \AD{≡} \AB{g} \AB{x}}.}
\agda{ν∘𝒟}
Moreover, the function \AF{𝒟} distributes over \AIC{[]} and \AF{\_⊙\_}:\footnote{In algebraic terms, argument-flipped \AF{𝒟} is a monoid homomorphism from lists to endofunctions.}
\agda{𝒟[]⊙}
These facts suggest that \AB{𝒟} can be computed one list element at a time via a more specialized operation, as is indeed the case:\footnote{Repeated \AF{δ} application is expressed here via a standard left fold\out{ \stdlibCitep{Data.List}}:
\agda{foldl}\vspace{-2ex}
In \AF{𝒟foldl}, {\AB{X} \AS = \AB{A} \AF ✶ \AS → \AB B}.
}
\agda{δ}
%\vspace{-4ex}
\agda{𝒟foldl}
Each use of \AF{δ} thus takes us one step closer to reducing general language membership to nullability.
%% In other words,%
%% \agda{ν∘foldlδ}
This simple observation is the semantic heart of the syntactic technique of \emph{derivatives of regular expressions} as used for efficient recognition of regular languages \citep{Brzozowski64}, later extended to parsing general context-free languages \citep{Might2010YaccID}.
Lemmas \AF{ν∘𝒟} and \AF{𝒟foldl} liberate this technique from the assumption that languages are represented \emph{symbolically}, by some form of grammar (e.g., regular or context-free), inviting other representations as we will see in \secref{Automatic Differentiation}.

@@

Decidable parsing relies only on \AF{ν} and \AF{δ} (via {\AF{⟦\_⟧}}), which are defined in \figrefdef{symbolic-defs}{Symbolic differentiation (column-major/patterns)}{\agda{defs}}, as systematically derived from the lemmas of \figref{nu-delta-lemmas}.\out{\notefoot{Describe this derivation.}}
These definitions are meant to be read in ``column-major'' order, i.e., each column is one function definition.
Correctness is guaranteed by the types of \AF{ν} and \AF{δ} and so is proved automatically by type-checking \figref{symbolic-defs}.
(We could thus use \emph{any} definitions of \AF{ν} and \AF{δ} that type- and termination-check, although not many definitions would.)
